Including EW and mixed QCD-EW corrections into PDF fits

### 2020-06-22 paper/sec-pineappl.tex
Fix small Latex mistakes
Improve section 2

In this paper we introduce a new library called \textsc{PineAPPL}, which bins phase-space weights independently from the chosen PDF set.
The files produced in this way, generally called grids, can be used to quickly evaluate several observables and also to assess the impact of different PDF sets and their PDF uncertainties.
Finally they are the main theoretical input to a PDF determination.
In that sense \textsc{PineAPPL} is similar to \textsc{APPLgrid}~\cite{Carli:2010rw} and \textsc{fastNLO}~\cite{Kluge:2006xs,Wobisch:2011ij,Britzger:2012bs}, but it also understands EW, which are the main interest in this paper.
The following features distinguish it:
\begin{itemize}
\item Support for arbitrary fixed-order calculations in powers of $\alpha$, $\alphas$ or combinations thereof, e.g.\ in mixed QCD-EW corrections.
Furthermore, variations of the renormalisation and factorisation scale are supported, if needed.
For each needed combination of the couplings and logarithms of renormalisation and factorisation scale a separate subgrid is created (see section~\ref{sec:multi-coupling-expansion} for more details);
\item Support for all-order predictions coming from a resummation calculation or a photon-/parton-shower, which are important for some observables (see section~\ref{sec:results}),
\item A simple \textsc{C}-interface, with a wrapper for \textsc{Fortran} and \textsc{Python} (see appendix~\ref{app:example-program} for examples and documentation), which is needed for Monte Carlos and programs to read and write \textsc{PineAPPL} grids.
\textsc{PineAPPL} itself is written in Rust (see appendix~\ref{app:installation} for installation instructions).
\item Superior efficiency for datasets with more then 100 observables.
\end{itemize}
For \textsc{mg5\_aMC@NLO}~\cite{Alwall:2014hca,Frederix:2018nkq} the interfacing code is already implemented in the most recent version, which replaces the \textsc{aMCfast}~\cite{Bertone:2014zva} interface.
The interfacing code for other Monte Carlo generators should be easy to write, see appendix~\ref{app:example-program} for a small example program.
Finally, \textsc{PineAPPL} provides programs to convert \textsc{APPLgrids} and \textsc{fastNLO} tables to \textsc{PineAPPL} grids.

@@

Fixed-order partonic cross sections $a + b \to X$ supported by \textsc{PineAPPL} are expansions in powers of the strong coupling $\alphas$, the electromagnetic coupling $\alpha$, and the logarithms of $\xi_\mathrm{R} = \mu_\mathrm{R}^2 / Q^2$ and $\xi_\mathrm{F} = \mu_\mathrm{F}^2 / Q^2$,
\begin{equation}
\sigma_{ab} (x_1, x_2, \phi) = \sum_{k,l,m,n} \alphas^k \left( \xi_\mathrm{R} Q^2 \right) \alpha^l \log^m ( \xi_\mathrm{R} ) \log^n ( \xi_\mathrm{F} ) w_{ab}^{(k,l,m,n)} \left( x_1, x_2, Q^2, \phi \right) \text{.}
\label{eq:expansion}
\end{equation}
Above $Q^2$ denotes the Ellis--Sexton scale, which, if chosen dynamically, depends on the phase space $\phi$, $Q^2 = Q^2 (\phi)$.
If this is the case, then we assume the fractions $\xi_\mathrm{R}$ and $\xi_\mathrm{F}$ to be constants of phase space, however.
For the central scale choice, $\xi_\mathrm{R} = \xi_\mathrm{F} = 1$, the terms $m > 0$ and $n > 0$ are clearly not needed; they are only required if scale variations are desired.
For example, a seven-point scale variation sets the fractions to following values,
\begin{equation}
(\xi_\mathrm{R}, \xi_\mathrm{F}) \in \left\{ \bigl( 1, 1 \bigr), \bigl( \tfrac{1}{2}, \tfrac{1}{2} \bigr), \bigl( 2, 2 \bigr), \bigl( \tfrac{1}{2}, 1 \bigr), \bigl( 1, \tfrac{1}{2} \bigr), \bigl( 2, 1 \bigr), \bigl( 1, 2 \bigr) \right\} \text{,}
\end{equation}
and estimates the perturbative QCD uncertainty---no EW uncertainty is covered by this method---as the envelope of the cross section evaluated with the previous values.

@@

To give an example of eq.~\eqref{eq:expansion}, the following shows Drell--Yan lepton-pair production up to terms at NLO (with arguments $x_1$, $x_2$, and $\phi$ suppressed for the phase-space weights):
\begin{equation}
\begin{split}
\sigma_{ab}
    &= \alpha^2 w_{ab}^{(0,2,0,0)} \\
    &+ \alphas \left( \xi_\mathrm{R} Q^2 \right) \alpha^2 w_{ab}^{(1,2,0,0)} + \alphas \left( \xi_\mathrm{R} Q^2 \right) \log (\xi_\mathrm{F}) \alpha^2 w_{ab}^{(1,2,0,1)} (Q^2) \\
    &+ \alpha^3 w_{ab}^{(0,3,0,0)} + \log (\xi_\mathrm{F}) \alpha^3 w_{ab}^{(0,3,0,1)} (Q^2) \text{.}
\end{split}
\end{equation}
The first term with index $(0,2,0,0)$ is the LO term, the following line shows the NLO QCD correction, and the final line the NLO EW correction.
Note that for all terms the dependence on the renormalisation scale is only indirectly through $\alphas$, because 1) higher-order terms in $\alpha$ never generate a renormalisation scale dependence (in the $\alpha$ schemes that are valid according to section~\ref{sec:multi-coupling-expansion}) and 2) higher-order QCD corrections only introduce an explicit renormalisation scale dependence in counterterms with vertices with more than two gluons.
These terms, however, are not present in this process at NLO.
Both NLOs, however, have a contribution from a collinear counterterm that depends on the factorisation scale.

@@

Due to typical size of the couplings $\alphas^2 \sim \alpha$, it is naively expected that within the same order, i.e.\ for fixed $k + l$, terms with larger powers $\alphas^k$ dominate over those with smaller powers; however, in practise this naive expectation is not always true due to dynamic effects.
Some prominent examples are vector-boson scattering processes and TODO.

@@

A different choice partition the $(x_1, x_2, Q^2)$ space,
\begin{equation}
H = [x_\mathrm{min},x_\mathrm{max}]^2 \times [Q^2_\mathrm{min}, Q^2_\mathrm{max}] \ni (x_1, x_2, Q^2)
\end{equation}
along each axis into a small numbers of bins and to insert the phase-space weights $w$ into the corresponding discrete bin.
This method is straightforward, but given a finite number of bins this usually yields a bad approximation for the cross section.
This can be remedied by increasing the number of bins, which would also increase the space requirements of the grid.
There are, however, more refined methods, which keep the number of bins constant, while improving the quality of the approximation: Interpolation grids.

>>>>

In this paper we introduce a new library called \textsc{PineAPPL}, which bins phase-space weights independently from the chosen PDF set.
The resulting files, generally called grids, can be used to quickly convolute the observables with multiple PDF sets.
This has at least two applications:
\begin{itemize}
\item the study of the dependence of observables on PDF sets; e.g.\ to determine how the central predictions of observables change when changing the PDF set, the PDF uncertainties for different PDF set, etc.\ and
\item the determination of PDF sets themselves; the grids together with the corresponding experimental data are the main inputs for a PDF fit.
\end{itemize}
This puts \textsc{PineAPPL} in the same category than \textsc{APPLgrid}~\cite{Carli:2010rw} and \textsc{fastNLO}~\cite{Kluge:2006xs,Wobisch:2011ij,Britzger:2012bs}.
What sets \textsc{PineAPPL} apart from the other libraries is its support for electroweak (EW) predictions/corrections, which are the main interest in this paper.
More specifically, the following features are supported:
\begin{itemize}
\item Support for arbitrary fixed-order calculations in powers of $\alpha$, $\alphas$ or combinations thereof, in particular also mixed QCD-EW corrections.
Furthermore, variations of the renormalisation and factorisation scale are supported, if needed.
For each needed combination of the couplings and logarithms of renormalisation and factorisation scale a separate subgrid is created (see section~\ref{sec:multi-coupling-expansion} for more details);
%\item Support for all-order predictions coming from a resummation calculation or a photon-/parton-shower, which are important for some observables (see section~\ref{sec:results}),
\item A simple \textsc{C}-interface (see appendix~\ref{app:example-program} for examples and documentation), which is needed for Monte Carlos and programs to read and write \textsc{PineAPPL} grids.
\textsc{PineAPPL} itself is written in Rust (see appendix~\ref{app:installation} for installation instructions).
\item Another interface is the command-line program \texttt{pineappl}, which allows users of the generated grids to perform convolutions of the grid with PDF sets.
Additionally, it can print further information on how the corresponding luminosity function is constructed, which perturbative orders are stored, the relative size of each correction, the relative size of each partonic channel, etc.
%\item Possibility to import \textsc{APPLgrids} and \textsc{fastNLO} tables.
\item Superior efficiency and moderate memory requirements for differential distributions with more then 100 bins.
\end{itemize}
For \textsc{mg5\_aMC@NLO}~\cite{Alwall:2014hca,Frederix:2018nkq} the interfacing code is already implemented in a separate version, which is intended to replace the \textsc{aMCfast}~\cite{Bertone:2014zva} interface.
The interfacing code for other Monte Carlo generators should be easy to write, see appendix~\ref{app:example-program} for a small example program.

@@

Fixed-order partonic cross sections $a + b \to X$ supported by \textsc{PineAPPL} are expansions in powers of the strong coupling $\alphas$, the electromagnetic coupling $\alpha$, and the logarithms of $\xi_\mathrm{R} = \mu_\mathrm{R}^2 / Q^2$ and $\xi_\mathrm{F} = \mu_\mathrm{F}^2 / Q^2$,
\begin{equation}
\sigma_{ab} (x_1, x_2, \phi) = \sum_{k,l,m,n} \alphas^k \left( \xi_\mathrm{R} Q^2 \right) \alpha^l \log^m ( \xi_\mathrm{R} ) \log^n ( \xi_\mathrm{F} ) w_{ab}^{(k,l,m,n)} \left( x_1, x_2, Q^2, \phi \right) \text{.}
\label{eq:expansion}
\end{equation}
Above $Q^2$ denotes the Ellis--Sexton scale, which, if chosen dynamically, depends on the phase space $\phi$, $Q^2 = Q^2 (\phi)$.
If this is the case, then we assume the fractions $\xi_\mathrm{R}$ and $\xi_\mathrm{F}$ to be constants of phase space, however.
For the central scale choice, $\xi_\mathrm{R} = \xi_\mathrm{F} = 1$, the terms $m > 0$ and $n > 0$ are clearly not needed; they are only required if scale variations are desired.
For example, a seven-point scale variation sets the fractions to the following values,
\begin{equation}
(\xi_\mathrm{R}, \xi_\mathrm{F}) \in \left\{ \bigl( 1, 1 \bigr), \bigl( \tfrac{1}{2}, \tfrac{1}{2} \bigr), \bigl( 2, 2 \bigr), \bigl( \tfrac{1}{2}, 1 \bigr), \bigl( 1, \tfrac{1}{2} \bigr), \bigl( 2, 1 \bigr), \bigl( 1, 2 \bigr) \right\} \text{,}
\end{equation}
and estimates the perturbative QCD uncertainty---no EW uncertainty is covered by this method---as the envelope of the cross section evaluated with the previous values.

@@

To give an example of eq.~\eqref{eq:expansion}, the following shows Drell--Yan lepton-pair production up to terms at NLO (with arguments $x_1$, $x_2$, and $\phi$ suppressed for the phase-space weights):
\begin{equation}
\begin{split}
\sigma_{ab}
    &= \alpha^2 w_{ab}^{(0,2,0,0)} \\
    &+ \alphas \left( \xi_\mathrm{R} Q^2 \right) \alpha^2 w_{ab}^{(1,2,0,0)} + \alphas \left( \xi_\mathrm{R} Q^2 \right) \log (\xi_\mathrm{F}) \alpha^2 w_{ab}^{(1,2,0,1)} (Q^2) \\
    &+ \alpha^3 w_{ab}^{(0,3,0,0)} + \log (\xi_\mathrm{F}) \alpha^3 w_{ab}^{(0,3,0,1)} (Q^2) \text{.}
\end{split}
\end{equation}
The first term with index $(0,2,0,0)$ is the LO term, the following line shows the NLO QCD correction, and the final line the NLO EW correction.
Note that for all terms the dependence on the renormalisation scale is only indirectly through $\alphas$, because 1) higher-order terms in $\alpha$ never generate a renormalisation scale dependence (in the $\alpha$ schemes that are valid according to section~\ref{sec:multi-coupling-expansion}) and 2) higher-order QCD corrections only introduce an explicit renormalisation scale dependence in counterterms with vertices with more than two gluons.
These terms, however, are not present in this process at NLO\@.
Both NLOs, however, have a contribution from a collinear counterterm that depends on the factorisation scale.

@@

Due to the typical sizes of the couplings $\alphas^2 \sim \alpha$, it is naively expected that within the same order, i.e.\ for fixed $k + l$, terms with larger powers $\alphas^k$ dominate over those with smaller powers; however, in practise this naive expectation is not always true due to dynamic effects.
Some prominent examples are vector-boson scattering processes and TODO.

@@

A different choice partitions the $(x_1, x_2, Q^2)$ space,
\begin{equation}
H = [x_\mathrm{min},x_\mathrm{max}]^2 \times [Q^2_\mathrm{min}, Q^2_\mathrm{max}] \ni (x_1, x_2, Q^2)
\end{equation}
along each axis into a small numbers of bins and to insert the phase-space weights $w$ into the corresponding discrete bin.
This method is straightforward, but given a finite number of bins this usually yields a bad approximation for the cross section.
This can be remedied by increasing the number of bins, which would also increase the space requirements of the grid.
There are, however, more refined methods, which keep the number of bins constant, while improving the quality of the approximation: Interpolation grids.
